Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data

### 2021-07-02 Manuscript/00-main.Rmd
* Revise the notation section of the Method * Correct typo and grammar in intro
Revise Method section

---
title: Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data
author:
- name: Boyi Guo*
  num: a
- name: Byron C. Jaeger
  num: a 
- name: AKM Fazlur Rahman
  num: a   
- name: D. Leann Long
  num: a  
- name: Nengjun Yi
  num: a
address:
- num: a
  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
# - num: b
  # org: Department of Applied Things, University B, City B, Country B
corres: "*Corresponding author name, This is sample corresponding address. \\email{xxx@uab}"
presentaddress: This is sample for present address text this is sample for present address text
authormark: Guo et al.
articletype: Research article
received: 2021-01-01
revised: 2021-02-01
accepted: 2021-03-01
abstract: "Spike-and-slab priors have been extensively studied for variable selection, and recently been extended to function selection in generalized additive models (GAMs). Most of the previous works on fitting spike-and-slab GAM relies on Markov chain Monte Carlo (MCMC) algorithms to identify the promising subsets of functions. The computational burden of MCMC makes these models less feasible, especially in high-dimensional setting. Besides, most of the current spike-and-slab GAM consider an “all-in-all-out” approach for function selection, rendering less flexibility to answer if non-linear effects are necessary in the model. We propose Bayesian hierarchical generalized additive models for function selection and prediction. The proposed models employ spike-and-slab priors, mixture normal and mixture double exponential, for smooth and sparse solution. Two EM based algorithms, EM-IWLS and EM-Coordinate Descent, are developed for accelerated computation. Simulation studies and real data analysis demonstrated improved performance against the state-of-art models, mgcv and COSSO. The software implementation of the proposed methods is freely available via the R package BHAM."
keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent
bibliography: bibfile.bib
output: 
  rticles::sim_article:
    keep_tex: TRUE
---

\newcommand{\bs}[1]{\boldsymbol{#1}}

>>>>

---
title: Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data
author:
- name: Boyi Guo*
  num: a
- name: Byron C. Jaeger
  num: b 
- name: AKM Fazlur Rahman
  num: a   
- name: D. Leann Long
  num: a  
- name: Nengjun Yi
  num: a
address:
- num: a
  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
- num: b
  org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
corres: "*Corresponding author name, This is sample corresponding address. \\email{xxx@uab}"
presentaddress: This is sample for present address text this is sample for present address text
authormark: Guo et al.
articletype: Research article
received: 2021-01-01
revised: 2021-02-01
accepted: 2021-03-01
abstract: "Spike-and-slab priors have been extensively studied for variable selection, and recently been extended to function selection in generalized additive models (GAMs). Most of the previous works on fitting spike-and-slab GAM relies on Markov chain Monte Carlo (MCMC) algorithms to identify the promising subsets of functions. The computational burden of MCMC makes these models less feasible, especially in high-dimensional setting. Besides, most of the current spike-and-slab GAM consider an “all-in-all-out” approach for function selection, rendering less flexibility to answer if non-linear effects are necessary in the model. We propose Bayesian hierarchical generalized additive models for function selection and prediction. The proposed models employ spike-and-slab priors, mixture normal and mixture double exponential, for smooth and sparse solution. Two EM based algorithms, EM-IWLS and EM-Coordinate Descent, are developed for accelerated computation. Simulation studies and real data analysis demonstrated improved performance against the state-of-art models, mgcv and COSSO. The software implementation of the proposed methods is freely available via the R package BHAM."
keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent
bibliography: bibfile.bib
output: 
  rticles::sim_article:
    keep_tex: TRUE
---

\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tp}{\text{pen}}

### 2021-07-02 Manuscript/01-Intro.Rmd
Revise Method section

Semi-parametric regression models are a more appropriate alternative to the "black-box" models, thanks to their balancing between model flexibility and interpretability. Among those, generalized additive models (GAMs), proposed in the seminal work of @Hastie1987, grew to be one of the most utilized modeling tools. In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu. \phi)$, can be expressed as the summation of smooth functions, $f_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as 
$$
E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
$$
where $g^{-1}$ is the inverse function of a smooth monotonic link function. The smooth functions can take many forms and are estimated using a pseudo-weighted version of the backfitting algorithm [@Breiman1985]. Nevertheless, the classic GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis in biomedical studies.

>>>>

Semi-parametric regression models are a more appropriate alternative to the "black-box" models, thanks to their balancing between model flexibility and interpretability. Among those, generalized additive models (GAMs), proposed in the seminal work of @Hastie1987, grew to be one of the most utilized modeling tools. In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$^[TODO: explain $\phi$], can be expressed as the summation of smooth functions, $f_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as 
$$
E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
$$
where $g^{-1}$ is the inverse function of a smooth monotonic link function. The smooth functions can take many forms and are estimated using a pseudo-weighted version of the backfitting algorithm [@Breiman1985]. Nevertheless, the classic GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis in biomedical studies.

### 2021-07-02 Manuscript/02-Method.Rmd
Revise Method section

Following the GLM notation introduced in Section \label{introduction}, we have a generalized additive model 
$$
E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
$$
with the smoothing function $f_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e. the design matrix derived from the smoothing function $f_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$. The dimension of the design matrix depends on the choice of the smoothing function. For example, a cubic spline function with $k$ knots produces a $k+2$ elements column vector. Its corresponding smoothness penalty matrix is denoted as $\bs S_j$. In all, the GAMs can be formulated in a matrix form 
$$
E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
$$
where $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable, and the matrix transposing operation is denoted with a superscript $^t$. To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g. $f_j(x_j) = x_j$ with the smoothness penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.

To encourage proper smoothness of the smoothing functions, we adopt the idea of smoothing penalties from smoothing spline models. The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the penalized framework. However, the direct integration of smoothness penalties into sparsity penalties is not obvious. Given the smoothness penalty matrix is most often symmetric and positive semi-definite ^[ we limit the discussion to univariate smoothing function. However, the penalty matrix could possibly contain multiple matrices, e.g. tensor product smoother.], it can be integrated into the design matrix $\bs X$ via a re-parameterization. The re-parameterization follows @Marra2011, in which eigen-decomposed the penalty matrix $\bs S = \bs U \bs D \bs U^T$, where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order. To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the null space of the smooth function. We further write the orthonormal matrix $\bs U$, containing the eigenvectors as column, as an (row) augmentation of two sub-matrices $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\text{pen}}\end{bmatrix}$, where $\bs U^{\text{pen}}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e. the penalized space, and $\bs U^0$ contains the eigenvectors for the null space. We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\text{pen}} \end{bmatrix}$. An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ is scale-invariant ^[ I know this work for tensor spline (see @Wood2006), but I am not sure if it can be generalized to all smoothing functions]. For notation convenience, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the after re-parameterized basis function matrix for the $j$th variable hereafter. Similarly, the spline function can be expressed in the matrix form
$$
f_j(x_j) = f_j^0(x_j) + f_j^\tp (x_j) = {\beta_j^0}^T \bs X_j^0 + {\beta_j^\tp}^T \bs X_j^\tp ,
$$

@@

## Spike and Slab Spline Prior

For high-dimensional data analysis, a common Bayesian modeling approach is to impose a particular family of priors, the spike-and-slab (SS) priors on the coeffcient in a regression modelss. The term "spike-and-slab" was first coined in @Mitchell1988 describing the shape of the prior. It is a mixture distribution, comprised with a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
$$
 \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
$$

The most distinct feature of the SS priors family is that it is conditioned on a latent binary variable $\gamma_i$ that indicates whether the covariate $x_i$ is included in the model. There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$ and the slab density $f_{slab}(\cdot)$, for example @George1993; @George1997; @Chipman1996 for grouped variables; @Brown1998 for multivariate outcomes; @Ishwaran2005; @Clyde2004 and reference therein. Here we introduce the most popular spike-and-slab prior introduced in @George1993, as  stochastic search variable selection (SSVS). Given a regression problem
$$
Y = \bs X\beta+\epsilon, 
$$
we can assume the prior distribution of the regression parameters $\beta$ depends on a vector of binary variables $\gamma=(\gamma_1, \dots \gamma_p)^\prime$ as follows: given $\gamma_i$, $\beta_i$ are independent distributed as a mixture priors
$$
\beta_i | \gamma_i \sim (1-\gamma_i)N(0, \sigma^2 v_0) + \gamma_iN(0,\sigma^2 v_1).
$$
The prior of $\gamma$ is traditionally assumed to be identically independent Bernoulli distribution. Even the conjugate normal prior of $\bs \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bs \gamma$. @George1993 described a Gibbs sampling algorithm to approximate the posterior distribution of $\bs \gamma$ for variable selection. The same authors later compared SSVS with other Bayesian variable selection approaches, providing some empirical suggestions on the advantages and disadvantages of various formulations and estimation algorithms. [@George1997] Among the various problem SSVS addresses, high-dimensionality is not one of them, due to the intensive computation needs of MCMC algorithms. @George1997 suggested MCMC algorithms works well for medium size of predictors (p=25), which is not feasible for high-dimensional data, where the number of predictors easily exceeds 100.

## Spike and slab spline prior

A generalized spline model can be formulated as 
$$
y_i \sim EF(\mu_i, \phi), \quad g(\mu_i) = \sum\limits_{j=1}^p f_j(x_{ij}), \quad i = 1, \dots, n.
$$
The link function of the mean can be expressed in a matrix form
$$
g(\mu_i) = \sum\limits_{j=1}^p\left[f_j^0(x_{ij}) + f_j^{pen}(x_{ij})\right] = \sum\limits_{j=1}^p\left[{\beta_j^0}^T X_{ij}^0 + {\beta_j^{pen}}^T X_{ij}^{pen}\right]
 = \beta^T X_i,
$$
where
$$
\begin{aligned}
  X_i = \begin{bmatrix}X^0_{i1}\\X^{pen}_{i1}\\\vdots \\X^0_{ip}\\X^{pen}_{ip}\end{bmatrix}
& &
  \beta = \begin{bmatrix}\beta^0_1\\\beta^{pen}_{1}\\\vdots \\\beta^0_{p}\\\beta^{pen}_{p}\end{bmatrix}
\end{aligned}
$$

### Mixture-Normal Prior
Given the full deign matrix $X = \begin{pmatrix} X_1 & X_2 & \dots & X_n \end{pmatrix}^T$, the likelihood function of the generalized model can be written as
$$
p(y|X\beta, \phi) = \prod\limits^n_{i=1}p(y_i|\beta^T X_i, \phi).
$$

@@

### Mixture Double Exponential Prior
One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which renders problem when estimating the large effects. In contrast, distributions with heavier tails are preferred, for example $t$ distribution with small degree of freedom and double exponential distribution. Specifically, double exponential prior is equivalent to the LASSO penalty in the frequentist models, providing a sparse solution. Meanwhile, the estimation algorithm of LASSO penalty are extremely efficient, greatly reducing the computation needs. Recent developments of spike-and-slab priors include spike-and-slab LASSO prior, i.e. spike-and-slab mixture double exponential prior. We extend the spike-and-slab LASSO prior to high-dimensional GAMs, providing computationally efficiency and sparse solution to our proposed ssGAM-MN.

The set-up of ssGAM-LASSO model is similar to ssGAM-MN, except we replace the normal distribution with double exponential distribution. The prior distributions of the coefficients $\beta^0_j, \beta^{pen}_j$ for the $j$ predictor can be formulated as
$$
\begin{aligned}
  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1).
\end{aligned}
$$
The hyper-priors for $\gamma^0_j, \gamma^{pen}_j, \theta_j$ are kept the same as specified in ssGAM-MN.

## Alforithm for fitting ssGAMs

>>>>

Following the GLM notation introduced in Section \label{introduction}, we have a generalized additive model and its matrix form,
\begin{equation}\label{eq:gam}
E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
\end{equation}
with the smoothing function $f_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e. the design matrix derived from the smoothing function $f_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$. The dimension of the design matrix depends on the choice of the smoothing function. For example, a cubic spline function with $k$ knots produces a $k+2$ elements column vector. Its corresponding smoothness penalty matrix is denoted as $\bs S_j$. $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable, and the matrix transposing operation is denoted with a superscript $^T$. To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g. $f_j(x_j) = x_j$ with the smoothness penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.

To encourage proper smoothness of the smoothing functions, we adopt the idea of smoothing penalties from smoothing spline models. The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the penalized framework. However, the direct integration of smoothness penalties into sparsity penalties is not obvious. Given the smoothness penalty matrix is most often symmetric and positive semi-definite ^[ we limit the discussion to univariate smoothing function. However, the penalty matrix could possibly contain multiple matrices, e.g. tensor product smoother.], it can be integrated into the design matrix $\bs X$ via a re-parameterization. The re-parameterization follows @Marra2011, in which eigen-decomposed the penalty matrix $\bs S = \bs U \bs D \bs U^T$, where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order. To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the null space of the smooth function. We further write the orthonormal matrix $\bs U$, containing the eigenvectors as column, as an (row) augmentation of two sub-matrices $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\text{pen}}\end{bmatrix}$, where $\bs U^{\text{pen}}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e. the penalized space, and $\bs U^0$ contains the eigenvectors for the null space. We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\text{pen}} \end{bmatrix}$. An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ is scale-invariant ^[ I know this work for tensor spline (see @Wood2006), but I am not sure if it can be generalized to all smoothing functions]. For notation convenience, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the after re-parameterized basis function matrix for the $j$th variable hereafter. Similarly, the spline function can be expressed in the matrix form
$$
f_j(x_j) = f_j^0(x_j) + f_j^\tp (x_j) = \bs {\beta_j^0}^T \bs X_j^0 + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
$$
and the generalized additive model in Equation \ref{eq:gam} be
\begin{equation}\label{eq:gam-repa}
E(Y|\bs x) =  g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[a + \sum\limits^p_{j=1} ({\beta_j^0}^T \bs X_j^0 + {\beta_j^\tp}^T \bs X_j^\tp)\right].
\end{equation}

@@

## Spike-and-Slab Smooth Priors

The family of spike-and-slab (SS) priors regression models dominates Bayesian high-dimensional analysis for its utility in outcome prediction and variable selection. We defer to @Bai2021Review for an in-depth introduction to spike-and-slab priors. To summary, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
$$
 \beta|\gamma \sim (1-\gamma)f_{spike}(\beta) + \gamma f_{slab}(\beta).
$$
The most distinct feature of the SS priors is that it is conditioned on a latent binary variable $\gamma$ that indicates whether the variable $x$ is included in the model. There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$^[TODO: should I change to notation of $f$ to not confound with the smooth functions?] and the slab density $f_{slab}(\cdot)$, for example @George1993; @George1997; @Chipman1996 for grouped variables; @Brown1998 for multivariate outcomes; @Ishwaran2005; @Clyde2004 and reference therein.

<!-- Comments -->
<!-- Even the conjugate normal prior of $\bs \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bs \gamma$. @George1993 described a Gibbs sampling algorithm to approximate the posterior distribution of $\bs \gamma$ for variable selection. The same authors later compared SSVS with other Bayesian variable selection approaches, providing some empirical suggestions on the advantages and disadvantages of various formulations and estimation algorithms. [@George1997] Among the various problem SSVS addresses, high-dimensionality is not one of them, due to the intensive computation needs of MCMC algorithms. @George1997 suggested MCMC algorithms works well for medium size of predictors (p=25), which is not feasible for high-dimensional data, where the number of predictors easily exceeds 100. -->

The major criticism of early spike-and-slab priors receives is being computationally prohibitive. [@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a], ^[TODO: add more citation]. Particularly, the development of Spike-and-slab Lasso [@Rockova2018b; @Rockova2018] substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis ^[TODO: add Dr. Yi's papers]. The spike-and-slab lasso prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
\begin{equation} \label{eq:ssl}
\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
\end{equation}
The SSL also mitigated the problem of EMVS where the weak signals are not shrink to zero, favored in high-dimensional data analysis. We noticed that @Bai2021 introduces a spike-and-slab group prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015]. @Bai2021 takes an "all-in-all-out" strategy for function selection.

### Spike-and-Slab Lasso Prior
We introduce a novel prior for GAMs, particularly for high-dimensional nonlinear effect modeling with bi-level selection. The proposed prior extends from the spike-and-slab lasso prior described in Equation \ref{eq:ssl}. Given the reparameterized deign matrix $\bs X_j = \begin{bmatrix} \bs X_j^0 : \bs X_j^\tp \end{bmatrix}^T$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j$. Specifically, we impose independent group priors on the null space coefficients and on the penalized space coefficients respectively, 
$$
\begin{aligned}
  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), 
\end{aligned}
$$
where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variables, indicating if the model includes the linear effect and the non-linear effect of the $j$th variable respectively. ; $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given. These scale parameters $s_0$ and $s_1$ can be treated as tuning parameters and optimized via cross-validation. A discussion of how to choose the scale parameters comes in Section \ref{selecting-optimal-scale-values}. To note, this prior differs from previous group spike-and-slab lasso priors[TODO: add Dr. Yi's citation], where the $\bs \beta^0_j$ and $\bs \beta^\tp_j$ shares the same indicator variables $\gamma_j^0$, $\gamma_j^\tp$ respectively. It is possible to relax the assumption such that each coefficient $\beta$ to have its own latent indicator $\gamma$, but at the cost ofcomplicating the bi-level function selection. Conversely, it is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and non-linear effect, i.e. $\gamma_j = \gamma^0_j = \gamma^{pen}_j$. This converges^[Is "becomes" better?] to the SB-GAM [@Bai2021].

The reparameterization introduced in Section \ref{bayesian-hierarchical-additive-models} grants the validity of the proposed prior. First of all, the smooth function bases are linear dependent to each other based on the definitions, and ask for extra attention. The eigeondecomposition remedies the problem. Hence our prior can be set to be conditionally independent. Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions of variables, where scaling is mandate.

The rest of the hierchical prior follows the traditional spike-and-slab lasso piror: we set up the hyper-prior of $\bs \gamma$, to allow local adaption of the shrinkage, using a bernoulli distribution. The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^{pen}_j$, shares the same probability parameter,
$$
\begin{aligned}
\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
\end{aligned}
$$

This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the null space and penalty space of the spline function. The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^{pen}$ decides the smoothness of the spline function at basis function level. Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^{pen}$ are independently distributed for analytic simplicity. We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
$$
\theta_j \sim Beta(a, b).
$$
The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience. For simplicity, we focus on a special case of beta distribution, uniform (0,1). When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases. Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL.

\newpage

### Other Priors

@@

One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which renders problem when estimating the large effects. In contrast, distributions with heavier tails are preferred, for example $t$ distribution with small degree of freedom and double exponential distribution. Specifically, double exponential prior is equivalent to the LASSO penalty in the frequentist models, providing a sparse solution.

## Algorithms for Fitting BHAMs

### 2021-07-02 Manuscript/05-Discussion.Rmd
Revise Method section

@@

>>>>

Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^{pen}$ are independently distributed for analytic simplicity. However, this set-up could possibility result in a situation that is not theoretically possible: the non-linear component is selected, but the linear component is not. This can be addressed analytically by forcing to include the linear component when non-linear component is included. Another possible solution is to impose an dependent structure of $\gamma_{j}^{pen}$ on $\gamma_{j^{0}}$, i.e. $\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j$.

@@

  <!-- Such prior set-ups ignore the null space of the smooth function, and entirely removes the variable out of the function. The "all-in-all-out" assumption makes differentiate linear and non-linear effect harder. In contrasts, our prior set up provides a more flexible solution, as the penalty for smoothness is locally adapative for the bases and hence retain the the null space if necessary.  -->

The "all-in-all-out" set-up was criticized for its inability to answering an important research question, if the effects are non-linear or not. On the other hand, the model can be much more flexible by allowing each dimension of the penalized space to have an independent indicator, i.e. $\gamma^{pen}_{jk}$ of $\beta^{pen}_{jk}$ for $k = 1\dots, K_j$. This prior set-up generalized to the group spike-and-slab. The additional flexibility doesn't not necessarily grand the model neither theoretical justification nor analytic simplicity. The spline penalty assumes a common smooth shrinkage $\lambda_j$ applied to all coefficients of the penalized space, which contradicts to the model setup. Besides, having multiple indicator variables complicates the decision of if non-linear effects are necessary.
